---
title: "HW 1 Solutions"
date: "9/7/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("bayesplot","knitr","arm","ggplot2","rstanarm")

```

## 7.2 Fake-data simulation and regression: 
Simulate 100 data points from the linear model, y =  a + bx + error, with a = 5, b = 7, the values of x being sampled at random from a uniform  distribution on the range [0, 50], and errors that are normally distributed with mean 0 and standard deviation 3. 


### 7.2a 
Fit a regression line to these data and display the output. 

```{r}

a <- 5
b <- 7

x <- runif(100,0,50)
e <- rnorm(100,0,3)

y = a + b*x + e
fitmdl <- lm(y ~ x)

display(fitmdl)

```

### 7.2b 
Graph a scatterplot of the data and the regression line. 

```{r}

plot (x,y)

abline(fitmdl, col = "pink", lwd = 3)

```

### 7.2c 
Use the text function in R to add the formula of the fitted line to the graph. 

```{r}

fm <- round(coef(fitmdl), 2)
formula <- paste("y = ", fm[1], ifelse(sign(fm[2])==1, "+", "-"), abs(fm[2]), "x")

plot(x, y)
abline(fitmdl)

text(x = 30, y = 150, labels = formula)

```

## 7.3 Fake-data simulation and fitting the wrong model: 
Simulate 100 data points from the model,  y = a + bx + cx2 + error, with the values of x being sampled at random from a uniform  distribution on the range [0, 50], errors that are normally distributed with mean 0 and standard  deviation 3, and a, b, c chosen so that a scatterplot of the data shows a clear nonlinear curve. 


### 7.3 a
Fit a regression line stan_glm(y ~ x) to these data and display the output. 

```{r}

c <- 8

y = a + b*x + c*x^2 + e
fitmdl2 <- stan_glm(y~x, refresh = 0)
fitmdl2

```

### 7.3b
Graph a scatterplot of the data and the regression line. This is the best-fit linear regression.  What does “best-fit” mean in this context?

```{r}

plot(x, y)
abline(fitmdl2)

```
The line of best fit is a line that runs through a scatter plot of data points that best expresses the relationship between those points. In this context it will he expressing the general trend with which these values are progressing.

## 7.6 Formulating comparisons as regression models: 
Take the election forecasting model and simplify  it by creating a binary predictor defined as x = 0 if income growth is less than 2% and x = 1 if  income growth is more than 2%. 

```{r}

ghv_data_dir <- "https://raw.githubusercontent.com/avehtari/ROS-Examples/master/"
hibbs <- read.table (paste0(ghv_data_dir,"ElectionsEconomy/data/hibbs.dat"), header=T)
hibbs$x = ifelse( hibbs$growth>=2,1,0)

```

### 7.6a
Compute the difference in incumbent party’s vote share on average, comparing those two  groups of elections, and determine the standard error for this difference.

```{r}
v0 <- hibbs$x==0
m0 <- mean(hibbs[v0,"vote"])
v1 <- hibbs$x==0
m1 <- mean(hibbs[v1,"vote"])
d <- m1 - m0

n0 <- length(which(v0=="TRUE"))
n1 <- length(which(v1=="TRUE"))
se0 <- sd(hibbs[v0, "vote"])/sqrt(n0)
se1 <- sd(hibbs[v1, "vote"])/sqrt(n1)
se <- sqrt(se0^2 + se1^2)
d
se
```

### 7.6b
Regress incumbent party’s vote share on the binary predictor of income growth and check  that the resulting estimate and standard error are the same as above. 

```{r}
fitmdl1 <- lm(vote ~ x, data = hibbs)
summary(fitmdl1)
```

## 8.8 Comparing lm and stan_glm: 
Use simulated data to compare least squares estimation to default Bayesian regression: 

### 8.8a
Simulate 100 data points from the model, y = 2 + 3x + error, with predictors x drawn from  a uniform distribution from 0 to 20, and with independent errors drawn from the normal  distribution with mean 0 and standard deviation 5. Fit the regression of y on x data using  lm and stan_glm (using its default settings) and check that the two programs give nearly  identical results. 

```{r}

x <- runif(100, 0, 20)
e <- rnorm(100, 0, 5)
y = 2 + 3*x + e
fitmdl2 <- lm(y~x)
fitmdl2
stan2 <- stan_glm(y ~  x, data = data.frame(x,y), refresh = 0)
stan2

```

### 8.8b
Plot the simulated data and the two fitted regression lines. 

```{r}

plot(x,y)
abline(fitmdl2, col = "red", lwd = 5)
abline(stan2, col = "yellow", lwd = 3)

```

### 8.8c
Repeat the two steps above, but try to create conditions for your simulation so that lm and stan_glm give much different results. 

```{r,echo=F}

x <- runif(50, 0, 20)
e <- rnorm(50, 0, 5)
y = 2 + 3*x + e
fitmdl2 <- lm(y~x)
fitmdl2
stan2 <- stan_glm(y ~  x, data = data.frame(x,y), refresh = 0, iter = 20)
stan2

plot(x,y)
abline(fitmdl2, col = "red", lwd = 5)
abline(stan2, col = "green", lwd = 3)

```

## 10.1 Regression with interactions: 
Simulate 100 data points from the model, y = b0 + b1 x +  b2 z + b3 xz + error, with a continuous predictor x and a binary predictor z, coefficients  b = c(1, 2, -1, -2), and errors drawn independently from a normal distribution with mean 0  and standard deviation 3, as follows. For each data point i, first draw zi, equally likely to take  on the values 0 and 1. Then draw xi from a normal distribution with mean zi and standard  deviation 1. Then draw the error from its normal distribution and compute yi. 

### 10.1a
Display your simulated data as a graph of y vs. x, using dots and circles for the points with  z = 0 and 1, respectively. 

```{r}

b = c(1, 2, -1, -2)
error <- rnorm(100, 0, 3)
z <- rbinom(100, 1, 0.3)
x <- rnorm(100, z, 1)
y = b[1] + b[2]*x + b[3]*z + b[4]*x*z + error

shape = ifelse(z==1, 1, 20)
plot(x, y, pch = shape)

```

### 10.1b
Fit a regression predicting y from x and z with no interaction. Make a graph with the data  and two parallel lines showing the fitted model. 

```{r}

data = data.frame(x, y, z)
z0 <- filter(data, z == 0)
z1 <- filter(data, z == 1)
fitmdl3 <- stan_glm(y ~ z + x, data = data, refresh = 0)
fitmdl3

plot(y ~ x)

abline(fitmdl3$coef[1],fitmdl3$coef[2])
abline(fitmdl3$coef[1],fitmdl3$coef[3],fitmdl3$coef[2])

```

### 10.1c
Fit a regression predicting y from x, z, and their interaction. Make a graph with the data  and two lines showing the fitted model. 

```{r}

data = data.frame(x, y, z)
fitmdl4 <- stan_glm(y ~ z + x + z:x, data = data, refresh = 0)
fitmdl4

```


## 10.2 Regression with interactions: 
Here is the output from a fitted linear regression of outcome y on  pre-treatment predictor x, treatment indicator z, and their interaction: 

```{r}

table10.2 <- data.frame(median = c(1.2, 1.6, 2.7, 0.7),
                        MAD_SD = c(0.2, 0.4, 0.3, 0.5))

row.names(table10.2) <- c("Intercept", "x", "z", "x:z")
table10.2

```

### 10.2a
Write the equation of the estimated regression line of y on x for the treatment group and the  control group, and the equation of the estimated regression line of y on x for the control group. 

Treatment group
ytreatment = 1.2 + 1.6x

Control group
ycontrol = 3.9 + 2.3x


### 10.2b
Graph with pen on paper the two regression lines, assuming the values of x fall in the range  (0, 10). On this graph also include a scatterplot of data (using open circles for treated units  and dots for controls) that are consistent with the fitted model. 

```{r}



```

## 10.5 Regression modeling and prediction: 
The folder KidIQ contains a subset of the children and  mother data discussed earlier in the chapter. You have access to children’s test scores at age 3,  mother’s education, and the mother’s age at the time she gave birth for a sample of 400 children.


### 10.5a
Fit a regression of child test scores on mother’s age, display the data and fitted model,  check assumptions, and interpret the slope coefficient. Based on this analysis, when  do you recommend mothers should give birth? What are you assuming in making this recommendation? 

```{r}
kidiq <- read.csv(file = '/Users/zarawaheed/Documents/BostonUniversity/MA678/R/HW_1/kidiq.csv')

score_age <- stan_glm(kid_score ~ mom_age, data = kidiq, refresh = 0)
score_age

plot(kidiq$mom_age, kidiq$kid_score)
abline(score_age)

```
The slope coefficient is 0.69 and it shows that with every year increase in the mom's age, the child's IQ increases by 0.69. So the older the mother is, the higher will be the child's IQ. However, we need to keep in mind that it's safer for women to give birth until about 35 so it would be good to keep that in mind.


### 10.5b
Repeat this for a regression that further includes mother’s education, interpreting both slope  coefficients in this model. Have your conclusions about the timing of birth changed? 

```{r}

score_ae <- lm(kid_score ~ mom_age + mom_hs, data = kidiq)
score_ae
coefficients(score_ae)

```
The timing of birth should

### 10.5c
Now create an indicator variable reflecting whether the mother has completed high school or  not. Consider interactions between high school completion and mother’s age. Also create a  plot that shows the separate regression lines for each high school completion status group. 

```{r}
edu <- ifelse(kidiq$mom_hs==0, "red", "blue")

score_ae <- lm(kid_score ~ mom_age + mom_hs + mom_age:mom_hs, data = kidiq)

score_ae

plot(kidiq$mom_age, kidiq$kid_score, col = edu)
```

### 10.5d
Finally, fit a regression of child test scores on mother’s age and education level for the first  200 children and use this model to predict test scores for the next 200. Graphically display  comparisons of the predicted and actual scores for the final 200 children. 

```{r}

```

## 10.6 Regression models with interactions: 
The folder Beauty contains data (use file beauty.csv)  Beauty and  teaching  evaluations  from Hamermesh and Parker (2005) on student evaluations of instructors’ beauty and teaching  quality for several courses at the University of Texas. The teaching evaluations were conducted  at the end of the semester, and the beauty judgments were made later, by six students who had  not attended the classes and were not aware of the course evaluations. 

See also Felton, Mitchell, and Stinson (2003) for more on this topic. 

### 10.6a
Run a regression using beauty (the variable beauty) to predict course evaluations (eval),  adjusting for various other predictors. Graph the data and fitted model, and explain the  meaning of each of the coefficients along with the residual standard deviation. Plot the  residuals versus fitted values. 



```{r}
beautyy <- read.csv(file = "/Users/zarawaheed/Documents/BostonUniversity/MA678/R/HW_1/beautyy.csv")

mod_b <- stan_glm(eval ~ beauty, data = beautyy, refresh = 0)
mod_b
plot(beautyy$beauty, beautyy$eval)
res_mod_b <- resid(mod_b)
```

### 10.6b
Fit some other models, including beauty and also other predictors. Consider at least one  model with interactions. For each model, explain the meaning of each of its estimated  coefficients.

The first model, mod_bf, is using beauty and gender to predict course evaluations.
We can see from the coefficients that if the beauty score is 0 and the person is male, the evaluation will be 4.1. The coefficient of female, -0.2, is the difference between a male and females evaluation score, holding their beauty score constant. The Beauty coefficient shows that each with each point increase in the beauty score, the student evaluation score increases by 0.1.

The second model, mod_ba, is using beauty and age to predict course evaluations taking into account the interaction between beauty and age.
We can see from the coefficients that if the beauty score is 0 and the person is 0 years old, the evaluation will be 4. The Beauty coefficient shows that with each point increase in the beauty score, the student evaluation score decreases by 0.3. There is no impact of the age or the interaction between age and beauty in this case.

The third model, mod_bfa, is using beauty, gender and age to predict course evaluations.
We can see from the coefficients that if the beauty score is 0, the person is male, and the age is 0, the evaluation will be 4.2. The Beauty coefficient shows that each with each point increase in the beauty score, the student evaluation score increases by 0.1. The coefficient of female, -0.2, is the difference between a male and females evaluation score, holding their beauty score constant. Age has no effect of student evaluations.

The fourth model, mod_bfamlc, is using all the give data to predict course evaluations, meaning beauty, female, age, minority status, language, course ID are all being used.
This would not be a great model to use as we can't clearly see the impact of each individual variable.
We can see from the coefficients that if the beauty score is 0, the person is male, the person does not belong to a minority group, the person doesn not speak english, and they are not from the lower division, the evaluation will be 4.2. The coefficient of female, -0.2, is the difference between a male and females evaluation score, holding all other variables constant. The coefficient of minority, -0.1, is the difference between a minority's and non-minority's evaluation score, holding all other variables constant.The coefficient of nonenglish, -0.3, is the difference between a native english speaker's and a foreign language speaker's evaluation score, holding all other variables constant. The coefficient of lower, 0.1, is the difference between a lower division and upper division instructor's score, holding all other variables constant. The Beauty coefficient shows that each with each point increase in the beauty score, the student evaluation score increases by 0.1. Age and course credit seem to have no effect on student evaluations.

(Intercept)  4.2    0.1  
beauty       0.1    0.
female      -0.2    0.1  
age          0.0    0.0  
minority    -0.1    0.1  
nonenglish  -0.3    0.1  
lower        0.1    0.1  
course_id    0.0    0.0  
```{r}
mod_bf <- stan_glm(eval ~ beauty + female, data = beautyy, refresh = 0)
mod_bf
mod_ba <- stan_glm(eval ~ beauty + age + beauty:age, data = beautyy, refresh = 0)
mod_ba
mod_bfa <- stan_glm(eval ~ beauty + female + age, data = beautyy, refresh = 0)
mod_bfa
mod_bfamlc <- stan_glm(eval ~ beauty + female + age + minority + nonenglish + lower +course_id, data = beautyy, refresh = 0)
mod_bfamlc
```
The first model compares beauty score to gender. The intercept shows that the scoreIf the 

## 10.7 Predictive simulation for linear regression:
Take one of the models from the previous exercise.

### 10.7a
Instructor A is a 50-year-old woman who is a native English speaker and has a beauty score  of -1. Instructor B is a 60-year-old man who is a native English speaker and has a beauty  score of -0.5. Simulate 1000 random draws of the course evaluation rating of these two  instructors. In your simulation, use posterior_predict to account for the uncertainty in  the regression parameters as well as predictive uncertainty. 

```{r}
A <- data.frame(age = 50, female = 1, beauty = -1)
B <- data.frame(age =60, female = 0, beauty = -0.5)
A
B

simA <- posterior_predict(mod_ba, newdata = A, draws = 1000)
```

### 10.7b
Make a histogram of the difference between the course evaluations for A and B. What is the  probability that A will have a higher evaluation? 

```{r}
hist(A)
```

## 10.8 How many simulation draws: 
Take the model from Exercise 10.6 that predicts course evaluations  from beauty and other predictors. 

### 10.8a
Display and discuss the fitted model. Focus on the estimate and standard error for the  coefficient of beauty. 

```{r}
mod_ba
```

### 10.8b
Compute the median and mad sd of the posterior simulations of the coefficient of beauty,  and check that these are the same as the output from printing the fit. 

```{r}
mxmod_ba <- as.matrix((mod_ba))
medmod_ba <- apply(mxmod_ba, 2, median)
madmod_ba <- apply(mxmod_ba, 2, mad)
cbind(median = medmod_ba, mad_sd = madmod_ba)
```

### 10.8c
Fit again, this time setting iter = 1000 in your stan_glm call. Do this a few times in order  to get a sense of the simulation variability. 

```{r}
mod_ba2 <- stan_glm(eval ~ beauty + age + beauty:age, data = beautyy, iter = 1000, refresh = 0)
```

### 10.8d
Repeat the previous step, setting iter = 100 and then iter = 10. 

```{r}
mod_ba3 <- stan_glm(eval ~ beauty + age + beauty:age, data = beautyy, iter = 100, refresh = 0)

mod_ba4 <- stan_glm(eval ~ beauty + age + beauty:age, data = beautyy, iter = 10, refresh = 0)
```

### 10.8e
How many simulations were needed to give a good approximation to the mean and standard  error for the coefficient of beauty? 

I think if the model is set up correctly and all assumptions are clear, running a single test with the default iteration settings would give a good approximation of the mean and standard error for the coefficient of beauty.


